The Name of the Title is Hope

### 2024-09-25 submission_UIST/source/input/5_user_study_and_methods.tex
clean up

\section{User study \& Methods}
With this study, we wanted to find out whether the experience of agency can be preserved during physical action augmentation when the user's own brain signals serve as the control signal to a physical end effector. Therefore, we conducted a user study comparing three conditions of agency experience. With the first two conditions, INTENTION and EXTERNAL we queried users at the two edges of agency experience. In EXTERNAL participants had no intention to move and no control over their movement. On the other hand, in INTENTION, participants were acting as they would in their day-to-day lives, fully in control. We then investigated in a third condition where our AUGMENTED prototype sits on the spectrum between INTENTION and EXTERNAL, in other words, whether it preserved agency and to what degree. To answer our question, we employed a mixed methods approach including a psychometric test of intentional binding, one standardized question, and a qualitative (phenomenological) interview.

\subsection{Participants}
Nine participants (M = 30.00 years, SD = 3.81) were recruited from our local institution and through the institute's online participant pool. All participants were right-handed. Participants were compensated with course credit or 12 Euro per hour of study participation. Prior to their participation, they were informed of the nature of the experiment, recording, and anonymization procedures and signed a consent form. The experiment was approved by the ethics committee of [anonymized]. One participant had to be excluded from further data analyses due to significant deviations from the instructions in the execution of the task. Precisely, they did not initiate their movements at their own volition but rather immediately at the onset of each trial, thereby violating the task instruction of waiting 2-3 s after trial onset, see task description below for more detail.

\subsection{Apparatus}
The experimental setup, depicted in Figure~\ref{fig:setup}, comprised: (1) a 1-channel EMG device, (2) a 64-channel EEG system, (3) a medically-compliant EMS device connected to two electrodes worn on the forearm, and (4) a tablet to run the experiment and collect behavioral responses. To assist readers in replicating our experiment, we provide the necessary technical details, the complete source code for the experiment, the collected data, and the analysis scripts\footnote{[anonymized]}.
% TODO add links after anonymization is lifted.

@@

\indent\textbf{(2) EEG Recording.} EEG data was recorded from 64 actively amplified Ag/AgCl electrodes referenced to electrode FCz in an actiCap Snap cap using BrainAmp DC amplifiers from BrainProducts. Electrodes were placed according to the extended international 10–20 system \cite{Jasper1983-uw}. One electrode was placed under the right eye to provide additional information about eye movements (vEOG). After fitting the cap, all electrodes were filled with conductive gel to ensure proper conductivity, and electrode impedance was brought below 10k$\Omega$. EEG (and EMG) data were recorded with a sampling rate of 250 Hz.

@@

\subsection{Experimental Task, Design and Procedure}
Participants performed a simple tapping task in three conditions with 75 trials each: INTENTION, EXTERNAL, and AUGMENTED.

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/task_new.pdf}
    \caption{Interaction flow depicting one trial in our touchscreen tapping task.}
    \label{fig:progression}
\end{figure}

@@

\begin{figure*}[!h]
    \centering
    \includegraphics[width=\textwidth]{figures/method.png}
    \caption{Diagram summarizing: (a) signal processing and event labeling from muscle activity, (b) extraction of linear slope features from labeled brain signals and (c) real-time application of the BCI gating the muscle stimulation hardware.}
    \label{fig:method}
\end{figure*}

@@

@@

\subsubsection{Hypotheses Testing}
Prior to any statistical analyses of the intentional binding measure, outlier trials were rejected. We applied `extreme outlier removal' using Tukey's method~\cite{Tukey1949-sl}. Three time intervals were considered relevant to describe `regular' behavior across trials which would manifest by: (1) Tapping the screen in a reasonable interval after the fixation cross disappeared. An excessively short or long delay indicated that participants either tapped the screen prematurely by accident or they were checking in with the experimenter, respectively. (2) Providing a `reasonable' estimation in the intentional binding task. (3) The EMS stimulation leading to an \textit{immediate} screen tap. A long delay between the EMS trigger and the subsequent screen tap indicated that the stimulation was not strong enough in this trial to lead to muscle actuation resulting in a screen tap. Taken together, applying Tukey's method to each of these time windows and fusing the rejected trials led to the exclusion of 102 trials across all participants (M = 34.00, SD = 16.46) and a final data set of 1698 trials.

In line with the literature on intentional binding, we hypothesized that the time intervals should be underestimated for the INTENTION and the AUGMENTED conditions. This should not be the case for the EXTERNAL condition where users had no intention to move. Hence, when binding occurs, the intervals should be underestimated, hinting at a higher SoA. To test this, we fitted a linear mixed effects model with \textit{condition} (INTENTION, EXTERNAL, AUGMENTED) as a fixed effect and as a random effect, we considered the intercepts for participants ($time interval ~ condition + 1|participantID$). We obtained p-values of the coefficients by calculating likelihood-ratio tests. All parameters were estimated by maximum likelihood estimation~\cite{Pinheiro2006-bk}. Post hoc, we computed pairwise contrasts with the Tukey HSD (Honest Significant Differences), accounting for multiple companions with alpha correction~\cite{Lenth2020-xk}.

>>>>

\section{User study \& Methods}
In this paper, we present a prototype that uses a user's own brain signals as the control signal to a physical end effector. With a user study, we wanted to find out whether the experience of agency can be preserved during physical action augmentation when using our prototype.

In the user study, we compared three conditions of agency experience. With the first two conditions, INTENTION and EXTERNAL we queried users at the two edges of agency experience. In EXTERNAL, participants had no intention to move and no control over their movement. On the other hand, in INTENTION, participants were acting as they would in their day-to-day lives, fully in control. We then investigated in a third condition where our AUGMENTED prototype sits on the spectrum between INTENTION and EXTERNAL, in other words, whether it preserved agency and to what degree. To answer our question, we employed a mixed methods approach including a psychometric test of intentional binding, one standardized question, and a qualitative (phenomenological) exit interview.

To assist readers in replicating our prototype and experiment, we provide the necessary technical details, the complete source code for the BCI and the experiment, the collected data, as well as all the analysis scripts\footnote{[anonymized]}

\subsection{Participants}
Eleven participants (M = 29.9 years, SD = 4) were recruited from our local institution and through the institute's online participant pool. Participants were compensated with course credit or 12 Euro per hour of study participation. Prior to their participation, they were informed of the nature of the experiment, recording, and anonymization procedures and signed a consent form. The experiment was approved by the ethics committee of [anonymized]. One participant had to be excluded from further data analyses due to significant deviations from the instructions in the execution of the task. Precisely, they did not initiate their movements at their own volition but rather immediately at the onset of each trial, thereby violating the task instruction of waiting 2-3 s after trial onset, see task description below for more detail.
% All participants were right-handed

\subsection{Apparatus}
The experimental setup, depicted in Figure~\ref{fig:setup}, comprised: (1) a 1-channel Electromyography (EMG) device, (2) a 64-channel EEG system, (3) a medically-compliant EMS device connected to two electrodes worn on the forearm, and (4) a tablet to run the experiment and collect behavioral responses.

@@

\indent\textbf{(2) EEG Recording.} EEG data was recorded from 64 actively amplified Ag/AgCl electrodes referenced to electrode FCz in an actiCap Snap cap using BrainAmp DC amplifiers from BrainProducts. Electrodes were placed according to the extended international 10–20 system \cite{Jasper1983-uw}. One electrode was placed under the right eye to provide additional information about eye movements (vEOG). After fitting the cap, all electrodes were filled with conductive gel to ensure proper conductivity, and electrode impedance was brought below 10k$\Omega$ where possible. EEG (and EMG) data were recorded with a sampling rate of 250 Hz.

@@

\subsection{Experimental Task, Design and Procedure}
Participants performed a simple tapping task in three conditions with 75 trials each: INTENTION, EXTERNAL, and AUGMENTED

\begin{figure}[h]
    \centering
    \includegraphics[width=\columnwidth]{figures/task_new.pdf}
    \caption{Interaction flow depicting one trial in our touchscreen tapping task.}
    \label{fig:progression}
\end{figure}

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{figures/method.png}
    \caption{Diagram summarizing: (a) signal processing and event labeling from muscle activity, (b) extraction of linear slope features from labeled brain signals and (c) real-time application of the BCI gating the muscle stimulation hardware.}
    \label{fig:method}
\end{figure*}

@@

@@

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{figures/eeg_results_new.pdf}
    \caption{(a) Grand average event-related potential (ERP) at electrode Cz of pre-movement EEG data epochs preceding the last second before a movement, idle epochs of the same trial are plotted alongside but originate from a different time window, see section~\ref{eeg_methods}; (b) Bottom: Slope features for both idle and pre-movement classes at electrode Cz. Top: Scalp maps of slope values for both classes and all channels. (c): Mean and per participant ROC curves. Dotted line at 15 \% false positive rate indicate the selected threshold for the real-time application.}
    \label{fig:EEG_results}
\end{figure*}

@@

\subsubsection{Hypotheses Testing}
Prior to any statistical analyses of the intentional binding measure, outlier trials were rejected. We applied `extreme outlier removal' using Tukey's method~\cite{Tukey1949-sl} on three time intervals that well describe `regular' behavior across trials: (1) Tapping the screen in a reasonable interval after the fixation cross disappeared. An excessively short or long delay indicated that participants either tapped the screen prematurely by accident or they were checking in with the experimenter, respectively. (2) Providing a `reasonable' estimation in the intentional binding task. (3) The EMS stimulation leading to an \textit{immediate} screen tap. A long delay between the EMS trigger and the subsequent screen tap indicated that the stimulation was not strong enough in this trial to lead to muscle actuation resulting in a screen tap. Taken together, applying Tukey's method to each of these time windows and fusing the rejected trials led to the exclusion of 110 trials across all participants (M = 12.2, SD = 9.1). % and a final data set of 2140 trials.

In line with the literature on intentional binding, we hypothesized that the time intervals should be underestimated for the INTENTION and the AUGMENTED conditions. This should not be the case for the EXTERNAL condition where users had no intention to move. Hence, when binding occurs, the intervals should be underestimated, hinting at a higher SoA.

To test this, we fitted a linear mixed effects model with \textit{condition} (INTENTION, EXTERNAL, AUGMENTED) as a fixed effect and `participandID' as a random effect. The model was specified as `time interval $\sim$ condition + (1|participantID)' and fit using the `lme4' package~\cite{Bates2015-bh}. We included all single trial data points in the analyses and did not aggregate before fitting the model as fitting the random effect `participantID' corrects for the correlation of the repeated measurements within participants. We obtained a test statistic by calculating likelihood-ratio tests comparing the full model as specified above against the null model `time interval $\sim$ 1 + (1|participandID)'. All parameters were estimated by maximum likelihood estimation~\cite{Pinheiro2006-bk}. We computed post-hoc pair-wise tests for `condition' corrected for multiple comparisons (Tukey method) using the emmeans package~\cite{Lenth2020-xk}.
